Draft Documents

### 2022-11-16 src/Selectively_remove_variables.Rmd
updated my part (selectively removing variables)

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
library(knitr)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(stats)
library(tidyverse)
library(e1071)
library(car)
library(glmnet)
library(corrplot)
```
We first read the 
```{r}
countries <- read.csv("../data/countries.csv")
```

```{r}
head(countries)
```

@@

#dataset without the red 
#removing reg as it is a bit too complex to deal for now, not relevant

#subset1 <- subset(countries, select = c("edu.total", "fdi", "gcf", "gdp.dflt", "hlth", "income", "lbr.part", "mil", "pop.gwth.rural", "pop.gwth.urban", "pov", "reg", "trade", "unemp", "year"))
subset0 <- subset(countries, select = c("edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))

subset1 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.gdlt = log(subset0$gdp.dflt))
#dataset
```

```{r}
#converting non-numeric arguments into factors

subset1$income <- factor(subset1$income)
subset1$income <- as.numeric(subset1$income)

```{r}
#Perform standardisation

(df.countries0 <- subset1)
scaler <- apply(X = df.countries0, MARGIN = 2, FUN = sd)

(df.countries <- as.data.frame(apply(df.countries0, 2, function(x) x/sd(x, na.rm = TRUE))))

@@

```{r}
#check for correlation

corrplot(cor(df.countries, use = "complete.obs"), 
         method = "number",
         type = "upper",
         tl.col = "black", 
         tl.srt = 30
         )
```

```{r}
set.seed(123)
index <- sort(sample(x = nrow(df.countries), size = nrow(df.countries) * 0.8))
train_1 <- subset1[index,]
test_1 <- subset1[-index,]

@@

```{r}
#Boxplot of the predictor variables
```

```{r}
```

```{r}
#Everything
lm_1 <- lm(data= train_1, formula = pov ~ .)
summary(lm_1)
```
```{r}
#remove hlth 
lm_2 <- lm(data= train_1, formula = pov ~ edu.total + gcf + fdi + gdp.dflt + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + year + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
summary(lm_2)
plot(lm_2, which = 5)
```

```{r}
#Remove fdi
lm_3 <- lm(data= train_1, formula = pov ~ edu.total + gcf + gdp.dflt + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + year + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
summary(lm_3)
plot(lm_3, which = 5)
```
```{r}
#remove year 
lm_4 <- lm(data= train_1, formula = pov ~ edu.total + gcf + gdp.dflt + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
summary(lm_4)
plot(lm_4, which = 5)
```
```{r}
#remove gdp.dflt 
lm_4 <- lm(data= train_1, formula = pov ~ edu.total + gcf + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc + lfdi + lgdp.pc + lgdp.gdlt)
summary(lm_4)
plot(lm_4, which = 5)
```
```{r}
lm_5 <- lm(data= train_1, formula = pov ~ edu.total + gcf + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc + lfdi + lgdp.pc)
summary(lm_5)
plot(lm_5, which = 5)
```
```{r}
#removing lfdi
lm_6 <- lm(data= train_1, formula = pov ~ edu.total + gcf + income + lbr.part + mil + pop.gwth.urban + pop.gwth.rural + trade + unemp + gdp.pc  + lgdp.pc)
summary(lm_6)
plot(lm_6, which = 4)
```
```{r}
#
data0 <- subset(df.countries, select = c("edu.total","gcf","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp","gdp.pc", 
                                        "lgdp.pc"))
data <- na.omit(data0)
train.x <- data.matrix(subset(data, select = -pov))
train.y = data.matrix(data$pov)

corrplot(cor(data, use = "complete.obs"), 
         method = "number",
         type = "upper",
         tl.col = "black", 
         tl.srt = 30
         )
```

```{r}
#Find the ridge regression model
set.seed(123)
rate <- 0.8
train.size <- round(nrow(data)*rate)
sample(sample(nrow(data), train.size))
training <- data[sample, ]
test <- data[-sample, ]

@@

@@

```{r}
train.x <- model.matrix(pov ~., data = training)[, -1]
train.y <- training$pov
test.x <- model.matrix(pov ~., data = test)[,-1]
test.y <- test$pov
```

```{r}
model_ridge_trial1 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.1)
t(coef(model_ridge_trial1))
```

```{r}
model_ridge_trial2 <- glmnet(train.x, train.y, alpha = 0, lambda = 0.5)
t(coef(model_ridge_trial2))
```

```{r}
model_ridge_trial3 <- glmnet(train.x, train.y, alpha = 0, lambda = 1)
t(coef(model_ridge_trial3))
```
```{r}
table_summary <- cbind(coef(model_ridge_trial1),
                       coef(model_ridge_trial2),
                       coef(model_ridge_trial3))

colnames(table_summary) <- c("lambda = 0.1",
                             "lambda = 0.5",
                             "lambda = 1")

table_summary
```

@@

```{r}
mlr_adj <- lm(pov ~., data = training)

summary(mlr_adj)
```

>>>>

```{r}
library(knitr)
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
library(stats)
library(tidyverse)
library(e1071)
library(car)
library(glmnet)
library(corrplot)
```

```{r}
countries <- read.csv("../data/countries.csv")
```

```{r}
head(countries)
```
This approach takes variables that contain less than 35% of missing data.The previous analysis showed that the relations between the predicted variable and gdp.pc, gdp.dflt, and fdi are better fitted to logarithmic models. Hence, we transformed the data with the mutate() function before building the machine learning model.
```{r}

subset0 <- subset(countries, select = c("edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))

subset1 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.dflt = log(subset0$gdp.dflt))

```
subset1 contains the variables below 35% of missing values. 
```{r}
subset1 <- subset(subset1, select = c("edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))

subset1 <- na.omit(subset1)
```
We also performed standardisation to reduce the variance of our machine learning models. Below are the functions used to aid our standardisation step. 
```{r}
sd0 <- function(vct) {
  if (!is.numeric(vct)) {
    return(NA)
  }

  return(sd(vct, na.rm = T))
}

std0 <- function(vct, scl) {
  if (!is.numeric(vct)) {
    return(vct)
  }

  return(vct/scl)
}

scale <- function(data) {
  unlist(lapply(data, sd0))
}

standardise <- function(data) {
  scaler <- scale(data)

  numCols <- which(unlist(lapply(data, is.numeric)))
  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
  fct <- data[,-numCols]
  return(cbind(fct, num))
}

@@

df.countries0 <- standardise(subset1)
df.countries <- df.countries0
```

@@

First, we split the data into training and testing data. 
```{r}
set.seed(123)
index <- sort(sample(x = nrow(df.countries), size = nrow(df.countries) * 0.8))
train_1 <- subset1[index,]
test_1 <- subset1[-index,]

@@

Our first attempt 
```{r}
#Everything
lm_1 <- lm(data= train_1, formula = pov ~ .)
summary(lm_1)
```
```{r}
#remove hlth 
lm_2 <- lm(data= train_1, formula = pov ~ .-lfdi - lgdp.dflt)
summary(lm_2)
plot(lm_2, which = 5)
```

```{r}
#Remove lgdp.dflt
lm_3 <- lm(data= train_1, formula = pov ~ . -lgdp.dflt)
summary(lm_3)
plot(lm_3, which = 5)
```
```{r}
#remove lfdi
lm_4 <- lm(data= train_1, formula = pov ~ . -hlth - lgdp.dflt - lfdi)
summary(lm_4)
plot(lm_4, which = 5)
```
Our model shows that among all the 13 predictor variables that we started with, only 10 of them showed significance in building the model.

We then evaluated the results of the model. 
```{r}
#evaluating lm_4
eval.metrics.linreg <- function(actual, predicted) {
  residual <- actual - predicted
  mse <- mean(residual ^ 2)
  mae <- mean(abs(residual))
  rmse <-  sqrt(mse)
  mape <- mean(abs(residual / actual)) * 100

  data.frame(
    MSE = mse,
    MAE = mae,
    RMSE = rmse,
    MAPE = mape
  )
}
```

```{r}
actual <- train_1$pov
predicted <- predict(lm_4, newdata = train_1)
eval.metrics.linreg(actual, predicted)
```

```{r}
#subsetting df.countries to only contain variables present in lm_7

data0 <- subset1
data <- na.omit(data0)
train.x <- data.matrix(subset(data, select = -pov))
train.y <- data.matrix(data$pov)

@@

```{r}
#Find the ridge regression model
set.seed(123)
rate <- 0.6
train.size <- round(nrow(data)*rate)
sample(sample(nrow(data), train.size))
training <- data[sample, ]
test <- data[-sample, ]

```{r}
training <- na.omit(training)
train.x <- model.matrix(pov ~., data = training)[,-1]
train.y <- training$pov
test.x <- model.matrix(pov ~., data = test)[,-1]
test.y <- test$pov
```

@@

@@

### 2022-11-16 src/draft.Rmd
updated my part (selectively removing variables)

for (indvar in logIvs) {
  print(
    ggplot(countries.num2 %>% 
           filter(variable == indvar),
         aes(x = log(value), y = pov)) +
    geom_point() +
    geom_smooth(formula = y ~ x, method = lm) +
    labs(title = paste0("pov ~ log(", indvar, ")"),
         x = paste0("log(", indvar, ")"),
         y = "pov")
  )
}
```

@@

## 3.1. Use Complete Cases (Imtiyaaz
###  3.1.1. Model Fitting
###  3.1.2. Assessment
###  3.1.3. Interpretation

## 3.2. Selectively remove variables with high missing rate (Callista)
###  3.2.1. Model Fitting
###  3.2.2. Assessment
###  3.2.3. Interpretation

## 3.3. Update the data set as we select variables (Isaac)
###  3.3.1. Model Fitting 
```{r}
df_count<-read.csv("../data/countries1.csv")
set.seed(10)
index <- sort(sample(x = nrow(df_count), size = nrow(df_count)*.8))
train <- df_count[index, ]
test <- df_count[-index, ]
```

Through forward selection, we will do feature selection. We will stop the forward selection when the adjusted r-value>0.75.

Through iteration of different models. A linear model consisting of income as a predictor has the highest adjusted r-square value=0.6201, thus it is selected.

```{r cars}
summary(lm(formula = pov ~ income,data = train)) #0.6201
```

Since income as a predictor brought the highest adjusted r-value, we add another predictor to income to find the predictor added to income which produced the highest adjusted r-value. reg as a predictor has the highest adjusted r-value=0.7181

```{r}
summary(lm(formula = pov ~ reg+income,data = train)) #0.7181
```

The forward selection process continues and lgdp.pc is the new predictor added that produced highest adjusted r-square=0.7492
```{r}
summary(lm(formula = pov ~ lgdp.pc+income+reg,data = train)) #0.7492
```

The addition of predictor gdp.pc produces the highest adjusted r-square value above 0.75, adjusted r-square=0.7703
```{r}
summary(lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)) #0.7703
```

Fitting of multiple linear regression with predictor gdp.pc, income, reg and lgd.pc.
```{r}
lm1<-lm(formula = pov ~ gdp.pc+income+reg+lgdp.pc,data = train)
summary(lm1)
```

###  3.3.2. Assessment
Assessment of linear model lm1: 
the predicted model has some NA values which are removed. The error rates are too high, therefore, this method of variable selection may not be suitable
```{r}
eval.metrics.linreg <- function(actual, predicted) {
  residual <- actual - predicted
  mse <- mean(residual ^ 2)
  mae <- mean(abs(residual))
  rmse <-  sqrt(mse)
  mape <- mean(abs(residual / actual)) * 100

  data.frame(
    MSE = mse,
    MAE = mae,
    RMSE = rmse,
    MAPE = mape
  )
}
predicted<-predict(lm1,newdata=test)
actual<-test$pov
df23<-na.omit(data.frame(actual,predicted))
ac<-df23$actual
pr<-df23$predicted
eval.metrics.linreg(ac,pr)
```

###  3.3.3. Interpretation
gdp per capital, income class and region of the country can provide a model to predict the poverty headcount ratio in a country.

## 3.4. Imputation
**What is imputation?**  
Imputation is a technique to handle missing values by replacing missing data with substitute values. In our study, we focus on "item imputation", which means substituting for a component of a data point (i.e. a variable). The general idea is to take a value that preserve the property of the data (e.g., distribution, mean, standard deviation), and use other variables to predict the missing values (much like regression). Since we observed some correlations during descriptive analysis, we expect the imputation algorithm to work well.

@@

- `m = 3` imputed data sets
- `maxit = 20` iterations

@@

```{r}
set.seed(1984)
# split data
seeds <- countries1 %>% 
  group_by(country.code) %>% 
  filter(row_number() == 1)

plants <- countries1 %>% 
  group_by(country.code) %>% 
  filter(row_number() != 1)

isComplete <- which(complete.cases(plants))
idx <- sample(isComplete, replace = F, 0.2 * nrow(plants))

countries.train.4 <- plants[-idx,] %>% 
  rbind(seeds)
countries.test.4 <- plants[idx,]

# run the algorithm, or read from the file we generated as this is a very time-consuming process
# countries1.imputed <- mice(countries.train.4, m = 3, maxit = 20, method = meth)
# saveRDS(countries1.imputed, 'countries1.imputed.RData')
countries1.imputed <- readRDS('countries1.imputed.RData')
summary(countries1.imputed)
```
We can take a look into one of the imputed data sets.
```{r}
countries1.imputed.1 <- complete(countries1.imputed, 1)
summary(countries1.imputed.1)
sum(!complete.cases(countries1.imputed.1))
```
We found no missing cases as expected.

###  3.4.1. Model Fitting
#### A. Ordinary Linear Regression
We generated 3 sets of imputed (training) data. We can build separate models using each data set, and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details). With all the generated data sets.

@@

```{r}
# fit model on imputed set no. 1
summary(fit1[[1]])
```

@@

#### B. VIF

@@

#### B. Step-wise AIC
We can conduct a step-wise AIC variable selection. It is similar to the procedure we use in class, but based on a metric call AIC (Akaike Information Criterion), which is an estimator of prediction error and relative quality of statistical models. The lower AIC is, the better the model fits. \\

@@

Performance on train set and test set are slightly improved. However, we have simplified the model by removing some unnecessary terms. We can update our models.

@@

#### C. Regularization

@@

```{r}
optimalModel <- function(formula, train, test) {
  set.seed(1984)
  # standise and save number of train cases (to split later)
  std.data <- standardise(train, test)
  trainCases <- 1:nrow(train)
  # matrix-ise
  xs <- model.matrix(formula, std.data$final)[,-1]
  ys <- std.data$final$pov
  # split data
  train.x <- xs[trainCases,]
  train.y <- ys[trainCases]
  test.x <- xs[-trainCases,]
  test.y <- ys[-trainCases]
  # list of tested alpha, resolution = 0.1
  alphas <- seq(0, 1, 0.1)
  # build a bunch of models
  models <- lapply(alphas, \(a) cv.glmnet(train.x, train.y, type.measure = "mse", alpha = a))
  cv.error <- unlist(lapply(models, \(model) model$cvm[model$lambda == model$lambda.min]))
  # best model
  best.model.idx <- which.min(cv.error)
  # optimal alpha and lambda
  alpha.opt <- alphas[best.model.idx]
  lambda.opt <- models[[best.model.idx]]$lambda.min
  best.model <- glmnet(train.x, train.y, alpha = alpha.opt, lambda = lambda.opt)
  # predict for test data
  train.fit <- predict(best.model, train.x)
  test.fit <- predict(best.model, test.x)
  # evaluation - train
  k <- best.model$df
  train.r2 <- best.model$dev.ratio
  train.adjR2 <- adjR2(train.fit, train.y, k)
  train.resid <- train.fit - train.y
  train.mse <- mean(train.resid^2)
  train.rmse <- sqrt(train.mse)
  train.mae <- mean(abs(train.resid))
  train.mape <- mean(abs(train.resid / train.y))
  # evaluation - test
  test.r2 <- r2(test.fit, test.y)
  test.adjR2 <- adjR2(test.fit, test.y, k)
  test.resid <- test.fit - test.y
  test.mse <- mean(test.resid^2)
  test.rmse <- sqrt(test.mse)
  test.mae <- mean(abs(test.resid))
  test.mape <- mean(abs(test.resid / test.y))
  # final results
  results <- list(
    train = list(
      data = train,
      r2 = train.r2,
      adj.r2 = train.adjR2,
      fitted = train.fit,
      residuals = train.resid,
      mse  = train.mse,
      rmse = train.rmse,
      mae  = train.mae,
      mape = train.mape
    ),
    test = list(
      data = test,
      r2 = test.r2,
      adj.r2 = test.adjR2,
      fitted = test.fit,
      residuals = test.resid,
      mse  = test.mse,
      rmse = test.rmse,
      mae  = test.mae,
      mape = test.mape
    ),
    scaler = std.data$scaler,
    alpha = alpha.opt,
    lambda = lambda.opt,
    model = best.model
  )

@@

Build the optimal models using each imputed data set.

```{r}
# helper function to build from a specific imputed data
buildWith <- function(set, imputed, test, formula, fun) {
  train <- complete(imputed, set)
  model <- fun(formula, train, test)
  return(model)
}

@@

#### A. Homoscedasticity

@@

```{r}
for (mod in models) {
  resid <- mod$train$residuals
  fitted <- mod$train$fitted
  # plot(resid ~ pov, main = paste("Model", i))
  print(ggplot(mapping = aes(y = resid, x = fitted)) +
    geom_point() +
    geom_smooth(formula = y ~ x, method = loess))
}
```

@@

```{r}
lapply(models, \(m) {
  resid <- m$train$residuals
  # apply Goldfeld-Quandt test
  gqtest(formula = resid ~ 1, order.by = m$train$fitted)
})
```

There is heteroscedasticity in our models. [Generalized Least Squares With Unknown For of Variance](https://rpubs.com/cyobero/187387) is a possible remedy of this problem, but at the cost of model interpretation. The high variance in higher end of `pov` can be explained by the relative volatile economical and political climate in highly poor countries (one standard deviation from mean), leading to unstable effect of predictors.

#### B. Independence

@@

#### C. Normally Distributed Residuals

@@

#### D. Outliers

As elastic net models produced in the last section are not very fit for detecting outliers using R-support function. As they just removed dummy variables in `country.code`, we can using OLS to detect some potential outliers. They are the same models we produced after the step-AIC selection (`fit3` and `fit4`)

```{r}
for (mod in fit3) {
  plot(mod, which = 4)
}
```

Some obvious outliers are point 1122, 1327, 1395, 1396, 1424, 1474.

```{r}
out <- c(1122, 1327, 1395, 1396, 1424, 1474)
var <- c("country.code","country.name","year","pov","income","edu.total","hlth","mil","lbr.part","gcf","lfdi")
# See the original data instead of imputed data
countries.train.4[out,var]
```

These countries all appear very infrequently in our data set (only twice for South Sudan and St.Lucia). The gap between their appearance are quite large, too. Such reasons could have made the model struggle to predict their poverty states.

Guinea is an interesting point. Because the outlier data (87.9% in 1991) is quite different to the next data point (40.8% in 1994), a huge jump in just 3 years.

>>>>

for (indvar in logIvs) {
  print(
    ggplot(countries.num2 %>% 
           filter(variable == indvar),
         aes(x = log(value), y = pov)) +
    geom_point() +
    geom_smooth(method = lm) +
    labs(title = paste0("pov ~ log(", indvar, ")"),
         x = paste0("log(", indvar, ")"),
         y = "pov")
  )
}
```

@@

## 3.1. Use Complete Cases (To be done)
###  3.1.1. Model Fitting
###  3.1.2. Assessment
###  3.1.3. Interpretation

## 3.2. Selectively remove variables with high missing rate (To be done)
**1. About selectively removing variables with high missing rate**

Previous analysis showed that of all the variables in the countries.csv dataset, many of them had a high missing rate of above 35%, an arbitrary percentage that our group has chosen to filter the predictor variables. This section attempts to build a model on variables that have missing rates of below 35%.

The previous analysis showed that the relations between the predicted variable and gdp.pc, gdp.dflt, and fdi are better fitted to logarithmic models. Hence, we transformed the data with the mutate() function before building the machine learning model.
```{r}
subset0 <- subset(countries, select = c("edu.total", "fdi","gcf","gdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","gdp.pc" ))

subset1 <- subset0 %>% mutate(lfdi = log(subset0$fdi)) %>% mutate(lgdp.pc = log(subset0$gdp.pc)) %>% mutate(lgdp.dflt = log(subset0$gdp.dflt))

subset1 <- subset(subset1, select = c("edu.total", "lfdi","gcf","lgdp.dflt","hlth","income","lbr.part","mil","pop.gwth.rural","pop.gwth.urban","pov","trade", "unemp", "year","lgdp.pc"))

subset1 <- na.omit(subset1)
```
We also performed standardisation to reduce the variance of our machine learning models. Below are the functions used to aid our standardisation step.

```{r}
#helper function
sd0 <- function(vct) {
  if (!is.numeric(vct)) {
    return(NA)
  }

scale <- function(data) {
  unlist(lapply(data, sd0))
}

standardise <- function(data) {
  scaler <- scale(data)

  numCols <- which(unlist(lapply(data, is.numeric)))
  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
  fct <- data[,-numCols]
  return(cbind(fct, num))
}

df.countries0 <- standardise(subset1)
df.countries <- df.countries0
head(df.countries)
```

###  3.2.1. Model Fitting
##### A. Ordinary Linear Regression
First, we split the data into training and testing data. The training dataset contains 80% of the initial dataset, and the testing data contains 20% of the whole dataset. 
```{r}
set.seed(123)
index <- sort(sample(x = nrow(df.countries), size = nrow(df.countries) * 0.8))
train_1 <- subset1[index,]
test_1 <- subset1[-index,]

```
Our first model includes all the variables with missing rate lower than 35%. This model is assigned to `lm_1`.
```{r}
lm_1 <- lm(data= train_1, formula = pov ~ .)
summary(lm_1)
```
The summary of `lm_1` suggests that the model still contains predictor variables with pr(>|t|) > 0.05. We can imply that these variables do not provide a high significance to our model and are thus able to be removed. To avoid removing significant variables, we shall conduct the removal of the variables deliberately. In `lm_1`, we notice that the variable `lgdp.dflt` has a high Pr(>|t|). Hence, we remove `lgdp.dflt`.
```{r}
lm_2 <- lm(data= train_1, formula = pov ~ . -lgdp.dflt)
summary(lm_2)
```
Removing `lfdi`,
```{r}
lm_3 <- lm(data= train_1, formula = pov ~ .-lfdi - lgdp.dflt)
summary(lm_3)
```
Further removing `hlth`,
```{r}
lm_4 <- lm(data= train_1, formula = pov ~ . -hlth - lgdp.dflt - lfdi)
summary(lm_4)
plot(lm_4, which = 5)
```
Our model shows that among all the 13 predictor variables that we started with, only 10 of them showed significance in building the model.

##### B. VIF

##### c. Regularization
Manual multiple regression models are prone to standard error. Hence, Ridge regression and LASSO regression are also used to build our model. 
Later on, we will compare the results of all the models we generated.
```{r}
data0 <- subset1
data <- na.omit(data0)
train.x <- data.matrix(subset(data, select = -pov))
train.y <- data.matrix(data$pov)

```
Dividing into training and testing datasets.
```{r}
set.seed(123)
rate <- 0.6
train.size <- round(nrow(data)*rate)
sample(sample(nrow(data), train.size))
training <- data[sample, ]
test <- data[-sample, ]
```
Further dividing into `train.x`, `train.y`, `test.x`, and `test.y`.
```{r}
training <- na.omit(training)
train.x <- model.matrix(pov ~., data = training)[,-1]
train.y <- training$pov
test.x <- model.matrix(pov ~., data = test)[,-1]
test.y <- test$pov
```
Using the lambda value that yields the lowest MSE value. 
```{r}
cv_ridge <- cv.glmnet(train.x, train.y, alpha = 0)
glm_ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)

t(coef(glm_ridge))
```

```{r}
#train the ridge regression model using lambda.min

glm_Ridge <- glmnet(train.x, train.y, alpha = 0, lambda = cv_ridge$lambda.min)
t(coef(glm_Ridge))
```

Function to evaluate the results 
```{r}
eval_results <- function(fit, true) {
  actual <- data.matrix(true)
  SSE <- sum((actual - fit)^2)
  SST <- sum((actual - mean(actual))^2)
  R_square <- 1 - SSE/SST
  data.frame(
    MSE = MSE(fit, true),
    MAE = MAE(fit, true),
    RMSE = RMSE(fit, true),
    MAPE = MAPE(fit, true),
    R2  = R_square

  )
}
```

```{r}
fit <- predict(glm_Ridge, train.x)
true <- train.y
summary_Ridge_train <- eval_results(fit, true)

```

```{r}
fit <- predict(glm_Ridge, test.x)
true <- test.y
summary_Ridge_test <- eval_results(fit, true)

```

```{r}
summary_Ridge <- rbind(summary_Ridge_train[-5], summary_Ridge_test[-5])
rownames(summary_Ridge) <- c("Ridge_train", "Ridge_test")
knitr::kable(summary_Ridge, digits = 3)

```

###  3.2.2. Assessment
###  3.2.3. Interpretation

## 3.3. Update the data set as we select variables (To be done)
###  3.3.1. Model Fitting
###  3.3.2. Assessment
###  3.3.3. Interpretation

## 3.4. Imputation
**1. What is imputation**  
Imputation is a technique to handle missing values by replacing missing data with substitute values. In our study, we focus on "item imputation", which means substituting for a component of a data point (i.e. a variable). The general idea is to take a value that preserve the property of the data (e.g., distribution, mean, standard deviation), and use other variables to predict the missing values (much like regression). Since we observed some correlations during descriptive analysis, we expect the imputation algorithm to work well.

@@

- `m = 5` imputed data sets
- `maxit = 30` iterations

@@

```{r}
set.seed(1984)
# split data
isComplete <- which(complete.cases(countries1))
idx <- sample(isComplete, replace = F, 0.3 * nrow(countries1))
countries.train.4 <- countries1[-idx,]
countries.test.4 <- countries1[idx,]

# blank for no imputation, cart for variable requiring imputation
meth <- c(rep("", 4), "cart", "", rep("cart", 16))
names(meth) <- colnames(countries1)
meth

# run the algorithm
# countries1.imputed <- mice(countries1, m = 3, maxit = 20, method = meth)
# saveRDS(countries1.imputed, 'countries1.imputed.RData')
countries1.imputed <- readRDS('countries1.imputed.RData')
summary(countries1.imputed)
```
We can take a look into one of the imputated data sets.
```{r}
countries1.imputed.1 <- complete(countries1.imputed, 1)
summary(countries1.imputed.1)
sum(!complete.cases(countries1.imputed.1))
```
We found no missing cases as expected.

###  3.4.1. Model Fitting
##### A. Ordinary Linear Regression
We generated 3 sets of imputated (training) data. We can build separate models using each data set, and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details). With all the generated data sets.

@@

```{r}
# fit model on imputated set no. 1
summary(fit1[[1]])
```

@@

##### B. VIF

@@

##### B. Step-wise AIC
We can conduct a step-wise AIC variable selection. It is similar to the procedure we use in class, but based on a metric call AIC (Akaike Information Criterion), which is an estimator of prediction error and relative quality of statistical models. The lower AIC is, the better the model fits. \\

@@

Performance on train set and test set are not affected. However, we have simplified the model by removing some unnecessary terms. We can update our models.

@@

##### C. Regularization

@@

```{r}
optimalModel <- function(formula, train, test) {
  set.seed(1984)
  # standise and save number of train cases (to split later)
  std.data <- standardise(train, test)
  trainCases <- 1:nrow(train)
  # matrix-ise
  xs <- model.matrix(formula, std.data$final)[,-1]
  ys <- std.data$final$pov
  # split data
  train.x <- xs[trainCases,]
  train.y <- ys[trainCases]
  test.x <- xs[-trainCases,]
  test.y <- ys[-trainCases]
  # list of tested alpha, resolution = 0.1
  alphas <- seq(0, 1, 0.1)
  # build a bunch of models
  models <- lapply(alphas, \(a) cv.glmnet(train.x, train.y, type.measure = "mse", alpha = a))
  cv.error <- unlist(lapply(models, \(model) model$cvm[model$lambda == model$lambda.min]))
  # best model
  best.model.idx <- which.min(cv.error)
  # optimal alpha and lambda
  alpha.opt <- alphas[best.model.idx]
  lambda.opt <- models[[best.model.idx]]$lambda.min
  best.model <- glmnet(train.x, train.y, alpha = alpha.opt, lambda = lambda.opt)
  # predict for test data
  train.fit <- predict(best.model, train.x)
  test.fit <- predict(best.model, test.x)
  # evaluation - train
  k <- best.model$df
  train.r2 <- best.model$dev.ratio
  train.adjR2 <- adjR2(train.fit, train.y, k)
  train.resid <- train.fit - train.y
  train.mse <- mean(train.resid^2)
  train.rmse <- sqrt(train.mse)
  train.mae <- mean(abs(train.resid))
  train.mape <- mean(abs(train.resid / train.y))
  # evaluation - test
  test.r2 <- r2(test.fit, test.y)
  test.adjR2 <- adjR2(test.fit, test.y, k)
  test.resid <- test.fit - test.y
  test.mse <- mean(test.resid^2)
  test.rmse <- sqrt(test.mse)
  test.mae <- mean(abs(test.resid))
  test.mape <- mean(abs(test.resid / test.y))
  # final results
  results <- list(
    train = list(
      data = train,
      r2 = train.r2,
      adj.r2 = train.adjR2,
      residuals = train.resid,
      mse  = train.mse,
      rmse = train.rmse,
      mae  = train.mae,
      mape = train.mape
    ),
    test = list(
      data = test,
      r2 = test.r2,
      adj.r2 = test.adjR2,
      residuals = test.resid,
      mse  = test.mse,
      rmse = test.rmse,
      mae  = test.mae,
      mape = test.mape
    ),
    scaler = std.data$scaler,
    alpha = alpha.opt,
    lambda = lambda.opt,
    model = best.model
  )

@@

Build the optimal models using each imputated data set.

```{r}
# helper function to build from a specific imputated data
buildWith <- function(set, imputed, test, formula, fun) {
  train <- complete(imputed, set)
  model <- fun(formula, train, test)
  return(model)
}

@@

##### A. Homoscedasticity

@@

```{r}
for (mod in models) {
  resid <- mod$train$residuals
  fitted <- mod$train$data$pov + mod$train$residuals
  # plot(resid ~ pov, main = paste("Model", i))
  print(ggplot(mapping = aes(y = resid, x = fitted)) +
    geom_point() +
    geom_smooth(formula = y ~ x, method = loess))
}
```

@@

```{r}
lapply(models, \(m) {
  resid <- m$train$residuals
  # apply Goldfeld-Quandt test
  gqtest(formula = resid ~ 1, order.by = m$train$data$pov)
})
```

There is heteroscedasticity in our models. [Generalized Least Squares With Unknown For of Variance](https://rpubs.com/cyobero/187387) is a possible remedy of this problem, but at the cost of model interpretation. The high variance in higher end of `pov` can be explained by the relative volatile economics and politic climate in highly poor countries ($pov > 0.25$), leading to unstable effect of predictors.

##### B. Independence

@@

##### C. Normally Distributed Residuals

@@

##### D. Outliers

As elastic net model produced in
